---
title: "Incarceration and Inequality in the U.S."
author: "Paul Gialis"
date: '2022-11-20'
output: html_document
---

```{r echo=FALSE, eval=TRUE, message=FALSE}
source("../source/a4-helpers.R")
source("../source/analysis.R")
```

## Abstract 
[TODO: Include a short abstract (about 50 words)]

* Here's a test of a dynamic variable. Say **"`r q1_string`!!!** 
* Here's another test. A vector
of numbers: `r test_query2(num)`.

Recall that _{r echo=TRUE}_ can be used to display the code, 
which is very useful when developing and debugging a report. Here's an example: 

```{r echo=TRUE}
num <- 5
sum <- sum(test_query2(num))
print(sum)
```

## Introduction 
In this assignment, we analyzed a data set of incarceration in the U.S. by various different variables. Using this data, it is obvious that there is a clear distinction of incarceration rates based on race, and certain states present this trend stronger than others. Through this assignment, I was able to create data visualizations to expose these inequalities better than data could alone. As a white male, this allowed the facts to cement in my brain, even though I fall victim to privilege hazards. All U.S. citizens are stakeholders in this data set, especially people of color. 

## Data Summary 
[TODO]

## Growth of the U.S. Prison Population


## Growth of Prison Population by State


## Growth of Incarcerated Black and White Populations


## Map of Black Incarceration in the U.S.


_All the very best with your work!_

